The resurgence of the linear optics interferometer --- recent advances \& applications

### 2017-05-17 (PR final edits) linear_optics_review/paper.tex
peter final edits

The difficulty in producing nonlinear operations required for universal LOQC led researchers to seek out alternative ways for demonstrating its power in the near term. {\sc Bosonsampling} is the result of such an effort. It is a restricted model of non-universal quantum computation \cite{bib:Aaronson11} that constitutes sampling the photocounts from the probability distribution of identical bosons scattered by a passive linear optical interferometer. While not universal, the {\sc bosonsampling} scheme is strongly believed to implement a classically hard task. Consider such a circuit of $m$ input (and output) modes that is injected with $n$ indistinguishable single photons such that $m\sim \mathcal{O}(n^2)$. Then, the boson sampling task consists of generating a sample from the probability distribution of single-photon measurements at the output of the circuit. The probability of detecting a $s_j$ photons at the $j$th output mode is
\begin{align}
p(s_1,\ldots,s_m)=\frac{{\rm Per}(U_S)}{s_1!\ldots s_n!} \ ,	
\end{align}
where ${\rm Per}$ denotes the permanent of a matrix, and $U_S$ is obtained from the interferometer unitary $U$ by keeping its first n columns and repeating its $j$th row $s_j$ times. The appearance of the permanent, which is of \#P complexity, in this statistics contributes to the hardness of the boson sampling problem. It is also known to be as hard as stimulating the short time evolution of a certain Hamiltonian\cite{bib:Peropadre17}.

@@

 Owing to the simplicity of {\sc BosonSampling} and its overarching implications for computer science, multiple experiments were reported shortly after its theoretical conception \cite{bib:Tillmann13,bib:Spring13, bib:Broome13,bib:Crespi13}. A recent experiment showed a {\sc bosonsampling} machine beating an early computer \cite{bib:WangHe16}.
{\sc BosonSampling} has piqued so much interest, that a complete discussion would make up its own review, so we shall only outline some main developments here. One variant, known as scattershot {\sc bosonsampling}, proposes using pumped SPDC heralded photon sources to increase the rate of generating input photons \cite{bib:Lund14,bib:Bentivegna15}. Others varied the source input to cat-states \cite{bib:Rohde15} and thermal states \cite{bib:Tamma14_thermal} . An exciting modification allows the determination of molecular vibronic spectra \cite{bib:Huh15}, a task that goes beyond demonstrating a computationally hard task.

>>>>

The difficulty in producing nonlinear operations required for universal LOQC led researchers to seek out alternative ways for demonstrating its power in the near term. {\sc Bosonsampling} is the result of such an effort. It is a restricted model of non-universal quantum computation \cite{bib:Aaronson11} that constitutes sampling the photocounts from the probability distribution of identical bosons scattered by a passive linear optical interferometer. While not universal, the {\sc bosonsampling} scheme is strongly believed to implement a classically hard task. Consider such a circuit of $m$ input (and output) modes that is injected with $n$ indistinguishable single photons such that $m\sim \mathcal{O}(n^2)$. Then, the boson sampling task consists of generating a sample from the probability distribution of single-photon measurements at the output of the circuit. The probability of detecting a $s_j$ photons at the $j$th output mode is
\begin{align}
p(s_1,\ldots,s_m)=\frac{|{\rm Per}(U_S)|^2}{s_1!\ldots s_n!} \ ,	
\end{align}
where ${\rm Per}()$ denotes the permanent of a matrix, $S=\{s_1,\dots,s_m\}$ denotes an output photon-number configuration, and $U_S$ is obtained from the interferometer unitary $U$ by keeping its first n columns and repeating its $j$th row $s_j$ times. The appearance of the permanent, which is \#P-hard in general (a complexity class even harder than NP-hard), in this statistics contributes to the hardness of the boson sampling problem. Note that the number of configurations $|S|=\binom{n+m-1}{n}$ in the output superposition scales exponentially with $m$. For this reason, the boson sampler does not let us \emph{calculate} matrix permanents, as this would require knowing individual amplitudes with high precision, which would require an exponential number of samples. Rather, the boson sampler samples across a distribution of permanents without actually revealing any of them to the experimenter. Nonetheless, this is a classically intractable problem. It is also known that boson sampling is as hard as stimulating the short time evolution of a certain Hamiltonian\cite{bib:Peropadre17}.

@@

 Owing to the simplicity of {\sc BosonSampling} and its overarching implications for computer science, multiple experiments were reported shortly after its theoretical conception \cite{bib:Tillmann13,bib:Spring13, bib:Broome13,bib:Crespi13}. A recent experiment showed a {\sc bosonsampling} machine beating an early computer \cite{bib:WangHe16}.
{\sc BosonSampling} has piqued so much interest, that a complete discussion would make up its own review, so we shall only outline some main developments here. One variant, known as scattershot {\sc bosonsampling}, proposes using pumped SPDC heralded photon sources to increase the rate of generating input photons \cite{bib:Lund14,bib:Bentivegna15}, by inputting an SPDC source into \emph{every} mode rather than just the first $n$, thereby boosting the probability of preparing the desired $n$ photons. Others varied the source input to cat-states \cite{bib:Rohde15} and thermal states \cite{bib:Tamma14_thermal} . An exciting modification allows the determination of molecular vibronic spectra \cite{bib:Huh15}, a task that goes beyond demonstrating a computationally hard task.
